---
title: "homework3"
format: html
name: Hunter Steinle
date: June 1, 2025
---

https://github.com/hiplikehunter/ENVS-193DS_homework-03

## Set Up

```{r}
library(tidyverse)
library(janitor)
library(here)
library(flextable)
library(readxl)
library(png)
library(grid)
library(dplyr)
library(gt)
#loading data
personaldata <- read_excel(here("data/personaldata.xlsx"))
```

## Problem 1

a. Data summarizing

I could compare and see if wind speeds or temperature is correlated with my biking pace. I imagine that windy or hot conditions could slow down my pace and effort. Categorically, I could group data by weekdays versus weekends and sunny versus cloudy weather and see if that affects my biking pace as well.

b. Visualization

```{r}

ggplot(data = personaldata, #gg plot using my personal data
       aes(x = WeatherType, #weather as indicator variable
           y = Pace, # biking pace as response variable
           color = WeatherType)) + # use this to assign groups with colors
  geom_boxplot() + # creates a boxplot for data
  geom_jitter(height = 0, # overlays data points in jitter plot, no jitter up or down
              width = 0.05) + # a little jitter side to side
  scale_color_manual(labels = c("Overcast", "Sunny"), # find which groups to assign colors
                     values = c("yellow3", "#38f246")) + # find which colors to use
  theme(panel.background = element_blank()) + # takes out gray background
      theme(panel.grid.major = element_line(color = "#000000")) + # puts in some lines 
    theme(panel.grid = element_line(size = .1)) + # changes size of lines
  labs(x = "Weather Condition", # creates labels for x axis
       y = "Pace (mph)") + # creates label for y axis
  labs(color = "Weather Condition") # Title of legend


```

c. Caption

Relationship between Weather Condition (Overcast & Sunny) and biking pace (mph).

d. Table Presentation

```{r}
summary_table <- personaldata |> # creates new object new summarize data
  group_by(WeatherType) |> # Groups data by weather type
  summarise(Count = n(), # creates a column for count
    Mean_Pace = round(mean(Pace, na.rm = TRUE), 2), # calculates the means for the groups
    SD_Pace = round(sd(Pace, na.rm = TRUE), 2)) # calculates the sd for the gorups


summary_table |> # uses summary data just calcualted
  gt() |> # creates data table
  tab_header(title = "Biking Pace by Weather Type") |> # creates title
  cols_label(WeatherType = "Weather", # column name for weather type/ indicator variable
    Count = "Number of Rides", # column name for number of rides
    Mean_Pace = "Mean Pace (mph)", # column name for mean pace
    SD_Pace = "SD Pace (mph)") |>  # column name for st dev
  fmt_number(columns = c(Mean_Pace, SD_Pace), # formats the columns for the mean and sd
    decimals = 1) # limits number of decimals to 1

```


## Problem 2

a. Describe in words what an affective visualization could look like for your personal data

My data is represented by two weather conditions, cloudy and sunny, and how that relationship affects with my biking pace. My affective visualization could be represented by a drawing of a bike, with the two wheels assigned to each indicator variable. Around the tires, other information such as a title and legend could fit within or on top of the bike frame. The summary data can be represented within the tires with colors or perhaps size of spokes.

b. Create a sketch (on paper) of your idea

```{r}
grid.raster(readPNG(here("affective_visualization_draft.png")))
```


c. Make a draft of your visualization

```{r}
grid.raster(readPNG(here("affective_visualization_draft.png")))
```

d. Write an artist statement

My affective visualization depicts my biking pace in accordance to weather type. The two wheels on the drawn bike represent the two weather types I recorded, and different length of spokes signal the different biking paces I recorded. I think for my final image, I will take a picture of my actual bike and superimpose my data driven representation on top of the wheels. I was inspired by a classmate who represented their data in circles and colors and connected the idea of using circles to the circular shape of my bike tires.


## Problem 3

a. Revisit and summarize

The paper is investigating the effects of the introduced mosquitofish on populations of stream breeding frogs and tadpoles. This paper includes an ANOVA test. The response variable is mean percent survival of mosquito larvae. The predictor variable is the type of experiment, whether in the lab or field and whether there is high or low density of mosquito larvae.

```{r}
grid.raster(readPNG(here("statisticalcritiqueplot.png")))
```


b. Visual clarity

The figure shows general trends of the data and includes error bars for the data. The axes are in logical positions with the three separate graphs stacked on top of each other.


c. Aesthetic clarity

Many lines of data points overlap each other and makes it hard to clearly distingish from the different groups of data. Also the font sizes within the figure are much smaller than that of the figure font.


d. Recommendations

I would include different colors for the different groups of data to further distinguish them along with the different shapes. I would also include the different treatments associated with the three graphs within the figure, rather than just listing them within the caption. I would stretch the figure vertically as well, since the data points and lines are smushed on top of each other. I would also change the symbols associated with each group of data. For example, I would have the mosquito larvae data have the same shape, but have them either colored in or not based on the presence of tadpoles or not. Similarly, I would do the same for Hyla regilla so that the data is groups logically due to species and treatment.

